---
title: "Data Analysis for Demand During Crises"
author: "Ryan Martin"
date: "June 2, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(tidy.opts=list(width.cutoff=80),tidy=TRUE)
knitr::opts_chunk$set(cache = TRUE)
# Original saved in C:\Users\mrya\Desktop\BNDS\Analysis\WithdrawalDepositCovidAnalysis
```

# Read In Data

```{r readindata, echo = FALSE}
# Reading in the data

library(pacman)
p_load(haven, tidyverse, ggplot2,sjlabelled, lubridate, 
  readxl, xtable, gapminder, kable, kableExtra, RColorBrewer,zoo, here)
p_load(astsa)
my_folder <- here("data")
my_tex_folder = here("LiabilitiesFigures")  

datliabl=read_csv(paste(my_folder, "NICMonthEndTimeSeriesfromLiabilities.csv" , sep="/"))


p_load(zoo, lubridate)
filter = dplyr::filter
datliabl = datliabl %>% filter(year(date)>= 1998)# %>% filter(year(date) <2021)
datliabl = datliabl %>% mutate(net_withdrawal = NIC - dplyr::lag(NIC))
datliabl = datliabl %>% mutate(year_mon = as.yearmon(date)) 

financial_events <- tibble(
  date = as_date(c("2020-03-12",
     "2008-09-01", "2009-09-01", #Lehman Bankruptcy to 2000 ARRA (American Recovery and Reinvestment Act) could also use March 6 2009 as nadar of down jones as end
     "2001-09-01", "2002-01-01",
     "1998-10-01", "2000-01-01", #Year 2000 Information and Readiness Disclosure Act to start of new year
     "1998-01-04", "1998-03-01" #Ice storm start is January 4th when formed. End was unclear. The power outage lasted several weeks for some places, or even several months. https://en.wikipedia.org/wiki/January_1998_North_American_ice_storm
     )),
  Event_Name = c("Pandemic start", 
    "GFC start", "GFC end",
    "Sept 11 start", "Sept 11 end",
    "Y2K start", "Y2K end",
    "Ice Storm start", "Ice Storm end"
))

datagg = datliabl
datagg_short = datliabl %>% filter(date <="2020-05-01")

```

# First Look at Time Series


```{r}

ts_dat = ts(datagg)
# View(ts_dat)
plot(ts_dat)
# it looks like net_withdrawal is more stable
# than weekly nic percent change
simple_ts = datagg %>% select(net_withdrawal)
simple_ts = ts(simple_ts)#ts(simple_ts, start = datagg$date[1], end=datagg$date[nrow(datagg)])
plot(simple_ts)

ggplot(datagg, aes(x = date, y = net_withdrawal)) + geom_line() + ggtitle("Monthly Net Withdrawals vs Time") + xlab("") + theme_classic() + ylab("Net Withdrawal (Millions CAD)") + xlab("date")
ggsave( paste("MonthlyNetWithdrawalsFullTS.jpg", sep = ""),
    path = paste(my_tex_folder, "monthly", sep = "/"))  
```

Dec 1999- Jan 2000 stand out as a large spike in net withdrawals at the monthly level. 2020-2021 stand out as times of an irregularly-persistent positive net withdrawals. 


# Time Series Analysis of Frequency

Looking at the periodogram, what can we conclude?




```{r}

#periodogram
p_load(astsa)
simple_ts = simple_ts[!is.na(simple_ts)]
dat.per = mvspec(simple_ts, log="n") #n.used=1200, roughly 23 years


year_harmonics = 1:6/12
year_multiples = as.character(1:6)#paste("Year Harmonic", as.character(1:12), sep = " ")
year_harmonic_tib = tibble(year_harmonics, year_multiples)

dat4plot= tibble(frequency = dat.per$freq, spectogram = dat.per$spec)
ggplot(dat4plot) + geom_line(aes(x=frequency, y =spectogram)) +
  geom_vline(xintercept = year_harmonics, color= year_multiples) +
  ggtitle("Periodogram of Net Withdrawals 1998-2020 with Colored Year Harmonics")
ggsave( paste("RawNoteWithdrawalPeriodogramWithColors.jpg", sep = ""),
    path = paste(my_tex_folder, "monthly", sep = "/"))


ggplot(dat4plot) + geom_line(aes(x=frequency, y =spectogram)) + ggtitle("Tapered Periodogram of Net Withdrawals 1998-2020")
ggsave( paste("RawNoteWithdrawalPeriodogramNoColors.jpg", sep = ""),
    path = paste(my_tex_folder, "monthly", sep = "/"))



# adding a taper
#periodogram
dat.per.taper = mvspec(simple_ts, taper=.1, log="no") #n.used=1200, roughly 23 years


dat4plottaper= tibble(frequency = dat.per.taper$freq, spectogram = dat.per.taper$spec)
ggplot(dat4plottaper) + geom_line(aes(x=frequency, y =spectogram)) +
  geom_vline(xintercept = year_harmonics, color= year_multiples) + ggtitle("Tapered Periodogram With Colored, 1-Year Harmonics")
ggsave( paste("RawNoteWithdrawalPeriodogram.jpg", sep = ""),
    path = paste(my_tex_folder, "monthly", sep = "/"))

dat4plottaper= tibble(frequency = dat.per.taper$freq, spectogram = dat.per.taper$spec)
ggplot(dat4plottaper) + geom_line(aes(x=frequency, y =spectogram))


#looking at logged version. not very helpful
dat.perlog = mvspec(simple_ts, log="yes") #n.used=1200, roughly 23 years


```



```{r}
p_load(broom)
datagg = datagg %>% mutate(year = year(date), month = month(date))
reg1_out = lm(data=datagg, formula=net_withdrawal ~ as.factor(month) + year + I(year^2))

logdat = datagg%>% mutate(log_net_lag_nic = log(NIC/lag(NIC)))
reg_log_out = lm(data=logdat, formula=log_net_lag_nic ~ month + year + I(year^2))
# plot(reg_log_out) #residuals for this seem much wakier
plot(reg1_out)
p_load(BETS)

plot(reg1_out, which=1:6)

######################
# Try forecasted vs true as nic
forecast_nic =  tibble(date=datagg$date[-1], fitted_net_withdrawal =  reg1_out$fitted.values) %>% mutate(fitted_nic = cumsum(fitted_net_withdrawal))

datagg=datagg %>% mutate(Year = as_character(year))
netwithdrawal_pred_1999= augment(reg1_out, newdata= datagg %>% filter(year==2019))
ggplot(netwithdrawal_pred_1999, aes(x = date, y = .fitted)) + geom_line() + geom_line(data=netwithdrawal_pred_1999,
  aes(x=date, y = net_withdrawal), color="red") + ylab("Net Withdrawal") + 
  ggtitle("Predicted Net Withdrawal in Red vs True for 1999")
ggsave( paste("PredictedvsTrueNetWithdrawal1999.jpg", sep = ""),
    path = paste(my_tex_folder, "monthly", sep = "/"))



#datagg = datagg %>% mutate(month = as.numeric(week))
ggplot(datagg, aes(x=month, y = net_withdrawal, color=Year ))+ 
  geom_point()  + 
  geom_line(data = netwithdrawal_pred_1999, aes(x = month, y = .fitted), size=1.5, color="black" ) + ylab("Net Withdrawal (Millions CAD)") +
  ggtitle("Seasonally Predicted 2019 vs True Net Withdrawals All Years")
ggsave( paste("PredictedvsTrueNetWithdrawalAllYears.jpg", sep = ""),
    path = paste(my_tex_folder, "monthly", sep = "/"))




true_and_forecast = left_join(datagg, forecast_nic)
options(dplyr.width = Inf)
gap = true_and_forecast$NIC[2] - true_and_forecast$fitted_nic[2]
true_and_forecast = true_and_forecast %>% mutate(
  fitted_nic_cor = fitted_nic + gap)
ggplot(true_and_forecast, aes(x=date, y = fitted_nic_cor)) + geom_point() + geom_point(aes(x=date, y = NIC), color="red") + 
  ggtitle("Seasonal Fit to Notes in Circulation Data: True in Red, Forecast in Black")
ggsave( paste("SeasonalFitToNIC.jpg", sep = ""),
    path = paste(my_tex_folder, "monthly", sep = "/"))

# It seems that it fit the outliers too well there in March 2020. I ought to adjust that a little.

#Maybe I should have predicted the NIC and then backed out
# the differences?

# Now, working on the residuals
# Obviously, don't interpret any of the coefficients too literally
# from here on out because already used some degrees of freedom
simple_ts_fit = ts(reg1_out$residuals)

#periodogram
dat.per = mvspec(simple_ts_fit, log="no") #n.used=1200, roughly 23 years


# adding a taper
#periodogram
dat.per.taper = mvspec(simple_ts_fit, taper=.1, log="no")
dat4plottaper= tibble(frequency = dat.per.taper$freq, spectogram = dat.per.taper$spec)
ggplot(dat4plottaper) + geom_line(aes(x=frequency, y =spectogram)) +
  geom_vline(xintercept = year_harmonics, color= year_multiples) + ggtitle("Tapered Periodogram With Colored, 1-Year Harmonics")



p_load(stargazer)
stargazer(reg1_out)
```


# Classic Time Series Fits

Note that these classic modeling paradigms do much better with the monthly data than the weekly data. 
```{r}
# Looking at SARIMA models of raw net withdrawal data
# recall its ARIMA(p,d,q) x (P,D,Q)_s
# Season (S) is 12 from above periodogram work
acf2(simple_ts_fit) # looks like just seasonal left
model1 = sarima(datagg$net_withdrawal,p=5,d=0,q=0,
    P=2,D=0,Q=0,S=12,no.constant=TRUE)
#plot(model1)

model2 = sarima(datagg$net_withdrawal,p=5,d=0,q=0,
    P=0,D=0,Q=2,S=12,no.constant=TRUE)
  
model3 = sarima(datagg$net_withdrawal,p=3,d=0,q= 2, P = 1, D = 0, Q = 1, S = 12)

# Model2 is pretty good

model4 = sarima(datagg$net_withdrawal,p=3,d=0,q=2,
    P=0,D=0,Q=2,S=12,no.constant=TRUE)



k=4
# Detecting large values in residuals
threshold = k*var(model2$degrees_of_freedom)
ts_out =arima(simple_ts_fit, order=c(3,0,2),include.mean = FALSE)
plot(ts_out$residuals)
acf2(ts_out$residuals)


# Detecting large values in residuals
threshold = k*sd(simple_ts_fit)
large_resid = which(abs(simple_ts_fit) > threshold)
datagg$date[large_resid]

# what about *persistent positive sequences?


resid_tab = tibble(date=datagg$date[-1],
  resid=as.numeric(simple_ts_fit))
resid_tab = resid_tab %>% mutate(fivelagsum =    
  resid+dplyr::lag(resid,n=1) +
  dplyr::lag(resid,n=2) + dplyr::lag(resid,n=3) +
    dplyr::lag(resid,n=4) + dplyr::lag(resid,n=5),
  fourlagsum = resid+dplyr::lag(resid,n=1) +
  dplyr::lag(resid,n=2) + dplyr::lag(resid,n=3) +
    dplyr::lag(resid,n=4),
  threelagsum = resid+dplyr::lag(resid,n=1) +
  dplyr::lag(resid,n=2) + dplyr::lag(resid,n=3),
  twolagsum = resid+dplyr::lag(resid,n=1) +
  dplyr::lag(resid,n=2),
  onelagsum = resid+dplyr::lag(resid,n=1) +
  dplyr::lag(resid,n=2))

my_paired_colors = brewer.pal(n = 12, "Paired")
quick_color = my_paired_colors[c(6,5,2,1,10,4,3,7,8)]


# here you really see 2000 and 2020 stand out
ggplot(resid_tab, aes(x=date, y = fivelagsum)) + geom_point() +
  ggtitle(" Cumulative Summation of Residual and Lags up to Five vs Date") +
  geom_vline(data= financial_events, 
    aes( xintercept = date, label = Event_Name, 
      color = Event_Name), wt = 3.5 ) + 
  scale_color_manual(values = quick_color) + ylab("")
ggsave( paste("FiveLagSum.jpg", sep = ""),
    path = paste(my_tex_folder, "monthly", sep = "/")) 

ggplot(resid_tab, aes(x=date, y = fourlagsum)) + geom_point() +
  ggtitle(" Cumulative Summation of Residual and Lags up to Fourth vs Date")
ggsave( paste("FourLagSum.jpg", sep = ""),
    path = paste(my_tex_folder, "monthly", sep = "/"))

ggplot(resid_tab, aes(x=date, y = threelagsum)) + geom_point()+
  ggtitle(" Cumulative Summation of Residuals and Lags up to Third vs Date")
ggsave( paste("ThreeLagSum.jpg", sep = ""),
    path = paste(my_tex_folder, "monthly", sep = "/"))

ggplot(resid_tab, aes(x=date, y = twolagsum)) + geom_point()+
  ggtitle(" Cumulative Summation of Residuals and Lag up to Second vs Date")
ggsave( paste("TwoLagSum.jpg", sep = ""),
    path = paste(my_tex_folder, "monthly", sep = "/"))


ggplot(resid_tab, aes(x=date, y = onelagsum)) + geom_point() +
  ggtitle(" Cumulative Summation of Residual and One Residual Lag vs Date")  +
  geom_vline(data= financial_events, 
    aes( xintercept = date, label = Event_Name, 
      color = Event_Name), wt = 3.5 ) + 
  scale_color_manual(values = quick_color) + ylab("")
ggsave( paste("OneLagSum.jpg", sep = ""),
    path = paste(my_tex_folder, "monthly", sep = "/"))



# here you really see 2000 and 2020 stand out
ggplot(resid_tab, aes(x=date, y = resid)) + geom_point() +
  ggtitle("Residuals After Controlling for Seasonal Effects vs Date") +
  geom_vline(data= financial_events, 
    aes( xintercept = date, label = Event_Name, 
      color = Event_Name), wt = 3.5 ) + 
  scale_color_manual(values = quick_color) + ylab("")
ggsave( paste("ResidualsAfterSeasonalAdj.jpg", sep = ""),
    path = paste(my_tex_folder, "monthly", sep = "/")) 

```


# Other Fits for Seasonal Behavior

Note that, since there is no "return to trend" yet in NIC, the nonparametric fit sees the maintained pandemic behavior as the trend. Should be careful on where we fit the data.



```{r}
 
p_load(lme4, splines, nlme, earth,jtools)
datagg = datagg %>% mutate(Month = as_factor(month))
datagg = datagg %>% drop_na()
fit1 <- earth(net_withdrawal ~ year + Month, data=datagg,pmethod="none")
summary(fit1)


my_knots = quantile(datagg$year, p = c(.5))
fit2 = lm(net_withdrawal~bs(year, knots=my_knots) + Month, data=datagg) 
summary(fit2)


my_forecast2 = stats::predict(fit1, newdata=datagg)
forecast_nic_spline =  tibble(date=datagg$date, fitted_net_withdrawal = my_forecast2) %>% mutate(fitted_nic = cumsum(fitted_net_withdrawal))


datagg = datagg %>% mutate(month = as.numeric(month))
datagg=datagg %>% mutate(Year = as_character(year))

####################
# Now with Knotted
#####################

month_names = paste("Month", as.character(2:12), sep = "")

gen_plots <- function(my_model, my_name) {
  
  if( class(my_model)=="earth") {
    temp = row.names(my_model$coefficients)
    omit_these_coefs = rep("name", length(temp) - 11)
    count = 1
    for (my_coef_name in temp) {
      if (!isTRUE(as.logical(grep(pattern="Month", my_coef_name)))) {
        omit_these_coefs[count] = my_coef_name
        count = count + 1
      }
    }
  } else {
    temp = names(my_model$coefficients)
    omit_these_coefs = rep("name", length(temp) - 11)
    count = 1
    for (my_coef_name in temp) {
      if (!isTRUE(as.logical(grep(pattern="Month", my_coef_name)))) {
        omit_these_coefs[count] = my_coef_name
        count = count + 1
      }
    }
  }
  if (class(my_model)!="earth") {
    # earth package does not play well with other model output packages. too bad
    # looks like just calls (tidy(my_model))...)
  plot1 = jtools::plot_summs(my_model, omit.coefs = omit_these_coefs,
    ci_level = .95) + ggtitle("Monthly Fixed Effects")
  plot1
  ggsave( paste("MonthlyCoefficientPlot", my_name, ".jpg", sep = ""),
      path = paste(my_tex_folder, "monthly", sep = "/"))
  } else {
  plot1 = ggplot(tibble(coef_names = names(
    my_model$coefficients[!(row.names(my_model$coefficients) %in% omit_these_coefs), ]),
    coef_values = my_model$coefficients[!(row.names(my_model$coefficients) %in% 
      omit_these_coefs), ]),
      aes(x = coef_names, y = coef_values)) + geom_point() + 
    ggtitle("Monthly Fixed Effects") + xlab("Month") + ylab("Millions CAD")
  plot1
  ggsave( paste("MonthlyCoefficientPlot", my_name, ".jpg", sep = ""),
      path = paste(my_tex_folder, "monthly", sep = "/"))
  }
  
  my_forecast = stats::predict(my_model, newdata=datagg)
  forecast_nic_func =  tibble(date=datagg$date, 
    fitted_net_withdrawal = my_forecast) %>% 
    mutate(fitted_nic = cumsum(fitted_net_withdrawal))
  
  netwithdrawal_pred_1999_spl= datagg %>% 
    mutate(.fitted = my_model$fitted.values) %>% 
    filter(year == 2019)
  
  netwithdrawal_pred_1999_spl = netwithdrawal_pred_1999_spl %>%
    mutate(.fitted = stats::predict(my_model, newdata=datagg %>% filter(year==2019)))
  plot2 = ggplot(netwithdrawal_pred_1999_spl, aes(x = date, y = .fitted)) + 
    geom_line() + geom_line(data=netwithdrawal_pred_1999_spl,
    aes(x=date, y = net_withdrawal), color="red") + ylab("Net Withdrawal") + 
    ggtitle("Predicted Net Withdrawal in Red vs True for 1999")
  plot2 
  ggsave( paste("PredictedvsTrueNetWithdrawal1999", my_name, ".jpg", sep = ""),
      path = paste(my_tex_folder, "monthly", sep = "/"))
  
  
  plot3 = ggplot(datagg, aes(x=month, y = net_withdrawal, color=Year ))+ 
    geom_point()  + 
    geom_line(data = netwithdrawal_pred_1999_spl, aes(x = month, y = .fitted),
      size=1.5, color="black" ) + ylab("Net Withdrawal (Millions CAD)") +
    ggtitle("Seasonally Predicted 2019 vs True Net Withdrawals All Years")
  plot3
  ggsave( paste("PredictedvsTrueNetWithdrawalAllYears", my_name, 
    ".jpg", sep = ""),
    path = paste(my_tex_folder, "monthly", sep = "/"))
  
  true_and_forecast_spl = left_join(datagg, forecast_nic_func)
  gap_spl2 = true_and_forecast_spl$NIC[1] - true_and_forecast_spl$fitted_nic[1]
  true_and_forecast_spl = true_and_forecast_spl %>% mutate(
    fitted_nic_cor = fitted_nic + gap)
  plot4 = ggplot(true_and_forecast_spl, aes(x=date, y = fitted_nic_cor)) + 
    geom_point() + geom_point(aes(x=date, y = NIC), color="red") +
    ggtitle("Seasonal Fit to Notes in Circulation Data: True in Red, Forecast in Black")
  plot4
  ggsave( paste("SeasonalFitToNICspl", my_name, ".jpg", sep = ""),
      path = paste(my_tex_folder, "monthly", sep = "/"))
  return(list(plot1, plot2, plot3, plot4))
}


myfirstlist = gen_plots(fit1,"earth_model") 
mysecondlist = gen_plots(fit2,"one_knot_splines")
myfirstlist[1]
myfirstlist[2]
myfirstlist[3]
mysecondlist[1]
mysecondlist[2]
mysecondlist[3]
mysescondlist[4]


```



__Something else to do is look at residuals^2 *or* the changes in NIC squared.__ I saw with the DJIA that the original series didn't exhibit much Autocorrelation but the square of the series did, which made it a good candidate for GARCH.

```{r}
acf2(simple_ts_fit^2)
Seasonal_Residuals=simple_ts_fit
acf2(Seasonal_Residuals,max.lag = 60)
Square_Of_Seasonal_Residuals = simple_ts_fit^2
acf2(Square_Of_Seasonal_Residuals)
# Looks pretty autocorrelated! I should fit a GARCH here just to see!
simple_ts_fit.sq = simple_ts_fit^2/(1e16)
simple_ts_fit_scale = simple_ts_fit/1e8
p_load(xts, fGarch)
# note that the arima modeling fails very poorly when numbers
# are large! hessian matrix fails. had to scale by 1e16 to 
# get it to evaluate. once scale, works well.
arima.fit.sq = arima(simple_ts_fit.sq + 0, order=c(3,0,0))
 #ar(1) and ar(3) are significant

out.garch <- garchFit(~ arma(1,0) + garch(3,1) , data=simple_ts_fit_scale, cond.dist="std") 
plot(out.garch)
summary(out.garch) # looks ok, but lots of NAs


# this one looks pretty good!
out.garch2 <- garchFit(~ arma(3,0) + garch(3,0) , data=simple_ts_fit_scale, cond.dist="std") 
plot(out.garch2) 
summary(out.garch2) # only issue is AR(3) not signif
 # and GARCh higher than one not significant. try simpler model



# this one looks pretty good!
out.garch2.5 <- garchFit(~ arma(2,0) + garch(1,0) , data=simple_ts_fit_scale, cond.dist="std") 
plot(out.garch2.5) 
summary(out.garch2.5) #looks ok


out.garch3 <- garchFit(~ arma(1,1) + garch(1,0) , data=simple_ts_fit_scale, cond.dist="std") #keep getting 
summary(out.garch3) #fit here much better! all terms highly signif.
plot(out.garch3) # looks about same as AR(3)


# Trying a heavier MA model
out.garch4 <- garchFit(~ arma(1,3) + garch(1,3) , data=simple_ts_fit_scale, cond.dist="std") #keep getting 
summary(out.garch4) # the heavier MA is not significant at all
  # for GARCh part. looks fine for rest. should reduce ma in
  # garch
plot(out.garch4)



# Trying a heavier MA model
out.garch5 <- garchFit(~ arma(2,3) + garch(2,0) , data=simple_ts_fit_scale, cond.dist="std") 
summary(out.garch5) #starting to get some errors in this
plot(out.garch5)


out.garch6.5 <- garchFit(~ arma(1,3) + garch(2,0) , data=simple_ts_fit_scale, cond.dist="std") 
summary(out.garch6) # garch(1,0) probably better
plot(out.garch6)


out.garch6 <- garchFit(~ arma(1,3) + garch(1,0) , data=simple_ts_fit_scale, cond.dist="std") 
summary(out.garch6) # garch(1,0)
plot(out.garch6)

out.garch8 <- garchFit(~ arma(1,5) + garch(1,0) , data=simple_ts_fit_scale, cond.dist="std") 
summary(out.garch8) # garch(1,0)
plot(out.garch8) # AIC and BIC back up


out.garch7 <- garchFit(~ arma(1,4) + garch(1,0) , data=simple_ts_fit_scale, cond.dist="std") 
summary(out.garch7) # garch(1,0)
plot(out.garch7)

# I would say overall, I like 2.5, 3, 6 oor 7 is best.
# AIC and BIC prefer 7. So, lets go with that.

# don't quite understand why access with dollar rather than 
my_garch_fit = tibble(date = datagg$date, conditional_sd =
    out.garch7@sigma.t, final_residuals = out.garch7@residuals, 
    conditional_var = out.garch7@h.t)


ggplot(my_garch_fit, aes(x=date, y = conditional_sd)) + geom_line() + ggtitle("GARCH Conditional Standard Deviation Estimates") + ylab("Conditional Standard Deviation")
ggsave( paste("GarchSDwithoutTimings.jpg", sep = ""),
    path = paste(my_tex_folder, "monthly", sep = "/"))



ggplot(my_garch_fit, aes(x=date, y = conditional_var)) + geom_line() + ggtitle("GARCH Conditional Variance Estimate") + ylab("Conditional Standard Deviation") 
ggsave( paste("GarchVarwithoutTimings.jpg", sep = ""),
  path = paste(my_tex_folder, "monthly", sep = "/")) 



#probably that spike on the residuals comes from the lack of persistence of the high values!
ggplot(my_garch_fit, aes(x=date, y = final_residuals)) + geom_line() + ggtitle("Final Residuals vs Date")



my_paired_colors = brewer.pal(n = 12, "Paired")
quick_color = my_paired_colors[c(6,5,2,1,10,4,3,7,8)]
ggplot(my_garch_fit, aes(x=date, y = conditional_sd)) + 
  geom_line() + ggtitle("GARCH Conditional Standard Deviation Estimates") +
  geom_vline(data= financial_events, 
    aes( xintercept = date, label = Event_Name, 
      color = Event_Name), wt = 3.5 ) + 
  scale_color_manual(values = quick_color) + ylab("Conditional Standard Deviation")
ggsave( paste("GarchSDwithTimings.jpg", sep = ""),
    path = paste(my_tex_folder, "monthly", sep = "/"))


ggplot(my_garch_fit, aes(x=date, y = conditional_var)) + geom_line() + ggtitle("GARCH Conditional Variance Estimate") + ylab("") + geom_vline(data= financial_events, 
    aes( xintercept = date, label = Event_Name, 
      color = Event_Name), wt = 3.5 ) + 
  scale_color_manual(values = quick_color) + ylab("Conditional Variance")
  
ggsave( paste("GarchVarwithTimings.jpg", sep = ""),
  path = paste(my_tex_folder, "monthly", sep = "/")) 


```


*Also, try the threshhold models. This data looks non-Gaussian and non-Linear. Therefore, a TARMA or SETARMA model may be appropriate

```{r}
library(pacman)
p_load(tsDyn)

```



# Generate Main Data and Source Functions

```{r}

#############################################
# Data
#############################################

year_pairs = tibble(
    start_year = c(1998, 1998, 2005, 2011, 2016),
    end_year = c(2020, 2004, 2010, 2015, 2020))


############################################
# Functions
############################################

plot_NIC_by_year <- function(start_year, end_year) {
  getPalette = colorRampPalette(brewer.pal(n=9, "Set1")) #can make 11?
  colorCount = end_year - start_year + 1
  pp = ggplot( datagg %>% filter(year>= start_year) %>% filter(year<=end_year) %>% mutate(Year = as_character(year)), aes(x = month, y = NIC/1e6, color = Year)) +
    scale_color_manual(values = getPalette(colorCount)) + 
    geom_point() + geom_line() + ylab("NIC (millions CAD)") +
    ggtitle(paste("Monthly NIC by year from ", start_year, " to ", 
      end_year, sep = "")
    )
}

save_by_year <- function(start_year, end_year) {
  ggsave( paste(
    "NICbyMonth", start_year, "to", end_year, ".jpg", sep = ""),
    path = paste(my_tex_folder, "monthly", sep = "/"))
}

plot_NIC_time_series <- function(start_year, end_year) {
  my_paired_colors = brewer.pal(n = 12, "Paired")
  financial_events_local = financial_events %>% filter(
    year(date)>= start_year) %>% filter(year(date)<=end_year)
  Event_count = nrow(financial_events_local)
  quick_color = my_paired_colors[c(6,5,2,1,10,4,3,7,8)]
  quick_color = quick_color[1:Event_count]
  pp = ggplot( datagg %>% filter(year>= start_year) %>% filter(year<=end_year), aes(x = date, y = NIC/1e6)) +
    geom_point() + geom_smooth(color="grey") + 
    ylab("NIC (Millions CAD)") + 
  ggtitle("NIC Over Time") +
  geom_vline(data= financial_events_local, aes( xintercept = date, label = Event_Name, color = Event_Name), wt = 3.5 ) + 
  scale_color_manual(values = quick_color) + ggtitle(
    paste("Monthly NIC from ", start_year, " to ", end_year, sep = "")
  )
}

save_time_series <- function(start_year, end_year){
    ggsave( paste(
    "NICtimeseries", start_year, "to",  end_year, ".jpg", sep = ""),
    path = paste(my_tex_folder, "monthly", sep = "/"))
}


generate_all_plots <- function(plot_generator) {
  year_pairs =.GlobalEnv$year_pairs 
    #get(year_pairs, envir = .GlobalEnv)
  force(year_pairs)
  
  # environment(plot_generator) = .GlobalEnv
   #my_plots = year_pairs %>% transmute(
   # my_plots= list(plot_generator(begin_year,final_year)))
  my_plot_list = list()
  for (i in 1:nrow(year_pairs)) {
    my_plot_list[[i]] = plot_generator(year_pairs$start_year[i],
      year_pairs$end_year[i])
  }
  my_plot_list
}

print_and_save_them_all = function(my_plot_list, plot_type="Forgot") {
  if (plot_type == "by_year") {
    for (counter in 1:length(my_plot_list)) {
      force(counter)
      print(my_plot_list[[counter]])
      save_by_year(year_pairs$start_year[counter],
        year_pairs$end_year[counter] )
    }  
  } else if (plot_type == "time_series") {
    for (counter in 1:length(my_plot_list)) {
      force(counter)
      print(my_plot_list[[counter]])
      save_time_series(year_pairs$start_year[counter],
        year_pairs$end_year[counter] )
    }  
    
  } else {
    print("Please provide either \"by_year\" or \"time_series\" as a plot type")
  }
}


plot_them_all = function(my_plot_list){
  for (each_plot in my_plot_list) plot(each_plot)
} 


#####################
# Generating Output
####################

by_year_plot = generate_all_plots(plot_NIC_time_series)
time_series_plot =  generate_all_plots(plot_NIC_by_year)


###########
# plot and save
print_and_save_them_all(by_year_plot, plot_type = "by_year")
print_and_save_them_all(time_series_plot, plot_type = "time_series")

# or, a quick view

###########
# generating plots without saving 
plot_them_all(by_year_plot)
plot_them_all(time_series_plot)



quick_plots = generate_all_plots(plot_NIC_by_year)
plot_them_all(quick_plots)


```

